---
title: "simulation_jeffrey"
author: "Anil Battalahalli Sreenath"
date: "12/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import stuff

```{r}
set.seed(84884)
library(expint)
library(remotes)
library(sigmoid)
```

## Accept-Rejection sampling to generate Power Law samples with Exponential Cutoff

```{r}
alpha <- 2.2
lambda <- 0.3
xmin <- 1.1

theta_true <- c(alpha, lambda, xmin)

c <-  lambda^(1-alpha)/(gammainc(1-alpha, lambda*xmin))
```

PDF of Power Law with Exponential Distribution

```{r}
fp <- function(x){
  if (x > xmin){
  return(c * ((x^ (-1 * alpha)) * (exp(-1 * lambda * x))))
  }
  else {
    return(0)
  }
}
```

Covering distribution

```{r}
foo <- function(xmin, alpha){
  u <- runif(1,0,1)
  return(xmin*((1-u)^(1/(1-alpha))))
}

generate <- function(n, xmin, alpha){
  return(as.numeric(replicate(n, foo(xmin, alpha), simplify=FALSE)))
}

getCumulative <- function(x, xmin, alpha){
  return(1-((x/xmin)^(1-alpha)))
}

covering <- function(i){
  return(generate(1, xmin, alpha))
}
```

```{r}
c_c <- ((alpha-1)/xmin) * (xmin^alpha)
covering_pdf <- function(x){
  if (x > xmin){
    return(((alpha-1)/xmin)*((x/xmin)^(-alpha)))
    }
  else{
    return(0)
  }
}
```

```{r}

M <- c*(exp(-lambda*(xmin)))/c_c
```

```{r}
compute_ratio <- function(u){
  return(fp(u)/(M*covering_pdf(u)))
}
```

Accept-Reject Sampling

```{r}
samples <- list()
for (i in 1:500){
  sample <- covering()
  r <- compute_ratio(sample)
  result <- rbinom(1,1,prob=r)
  if (result == 1){
    samples <- c(samples, sample)
  }
}
samplesPL <- as.numeric(samples)
length(samplesPL)
```

```{r}
saveRDS(samplesPL, "samplesPL_2o5_1o5_2o1_500samples.rds")
```

```{r}
plot(density(samplesPL))
```

## Bayesian Analysis

1.  Alpha - Jeffrey's prior
2.  Lambda - known
3.  xmin - known

```{r}
lambda <- theta_true[2]
xmin <- theta_true[3]

# From the known quantities
```

Gamma function

```{r}
G <- function(alpha){
  suppressWarnings(gammainc(1-alpha, lambda*xmin))
}
```

First derivative of the Gamma function

```{r}
Gp1 <- function(alpha){
  -1 * alpha * suppressWarnings(gammainc(-1*alpha, lambda*xmin))
}
```

Second derivative of the Gamma function

```{r}
Gp2 <- function(alpha){
  (alpha*(alpha+1)*suppressWarnings(gammainc(-1-alpha, lambda*xmin))) - suppressWarnings(gammainc(-1*alpha, lambda*xmin))
}
```

```{r}
prior_alpha <- function(alpha){
    j <- ((G(alpha)*Gp2(alpha))-(Gp1(alpha)^2))/(G(alpha)^2)
    if (is.na(j)){
      return(0)
    }
    if (j < 0){
      j <- j*-1
    } else{
      return(sqrt(j))
    }
}
```

```{r}
pllikelihood <- function(x, alpha){
  if (x > xmin){
    return(((x^(-alpha)) * exp(-lambda * x) * (lambda^(1 - alpha)))/suppressWarnings(gammainc(1-alpha, lambda * xmin)))
  }
  else {
    return(1)
  }
}
```

```{r}
loglikelihood <- function(data, alpha){
  lik <- lapply(data, pllikelihood,  alpha=alpha)
  blee <- lapply(lik, log)
  return(sum(as.numeric(blee)))
}
```

Posterior distribution known up to a constant

```{r}
logpostfunc <- function(data, alpha){
  alpha <- as.numeric(alpha)
  return(loglikelihood(data, alpha)+log(prior_alpha(alpha)))
}
```

Proposal

```{r}
proposal <- function(given){
  k <- as.numeric(given)
  return(rexp(1, k))
}
```

```{r}
proposal_pdf <- function(this, given){
  return(dexp(as.numeric(this), as.numeric(given)))
}
```

Metropolis-Hasting for Bayesian Inference

```{r}
reps <- 100
acc <- list()
params <- list()
pb = txtProgressBar(min = 1, max = reps, initial = 1)
for (v in 1:reps){
  nit <- 100
  accepted <- 0
  theta <- list()
  theta[1] <- proposal(2.1)
  for(i in 2:nit){
    theta_star <- proposal(theta[i-1])
    a = (logpostfunc(samplesPL, theta_star) + log(proposal_pdf(theta[i-1], 
                    theta_star)))-logpostfunc(samplesPL,theta[i-1]) - log(proposal_pdf(theta_star,theta[i-1]))
    a <- exp(a)
    if (is.nan(a)){
      theta[i] <-  theta[i-1]
    } else if (a > 1){
      theta[i] <-  theta_star
      accepted <- accepted+1
    } else{
      r = rbinom(1,1,a)
      if (r == 1){
        theta[i] <-  theta_star
        accepted <- accepted+1
      } else{
        theta[i] <-  theta[i-1]
      }
    }
  }
  acc[v] <- accepted/nit
  post_pdf <- lapply(theta, logpostfunc, data=samplesPL)
  theta_selected <- theta[which.max(post_pdf)]
  params[v] <- theta_selected
  setTxtProgressBar(pb,v)
}
```

```{r}
params <- as.numeric(params)
write.csv(data.frame(alpha=params), "case_1_params.csv")
d = density(params)
0
```

```{r}
plot(d, xlab = "alpha", ylab = "Density", main = "Density of alpha on the posterior")
abline(v=d$x[i], col="blue")
abline(v=alpha, col="red")
```

```{r}
mean(params)
```